---
output:
  pdf_document: default
  html_document: default
---

```{R}
library(dplyr)
library(reshape2)
library(ggplot2)
library(scales)
library(tidyr)
library(stringr)
library(knitr)
```
# Tornado Plot
### Cost Ranges

```{r}
# Grab data from CSV
data <- read.csv("total_cost_summary.csv")

# Make first column the index and remove first column
row.names(data) <- data[,1]
data<- data[, -1]

# Rearrange into columns of lower and upper value
C1_bounds <- data.frame(lower = data[c(6, 8, 1, 10, 4), 1], upper = data[c(7, 9, 3, 11, 5), 1])
C2_bounds <- data.frame(lower = data[c(6, 8, 1, 10, 4), 2], upper = data[c(7, 9, 3, 11, 5), 2])
C3_bounds <- data.frame(lower = data[c(6, 8, 1, 10, 4), 3], upper = data[c(7, 9, 3, 11, 5), 3])
C4_bounds <- data.frame(lower = data[c(6, 8, 1, 10, 4), 4], upper = data[c(7, 9, 3, 11, 5), 4])

# Allow Crowding
# Make dataset of first tornado plot, grab lower & upper estimates for variables
tornado_data_C1 <- C1_bounds %>% 
	mutate(param = c("AV cost/mile", 
	                 "Social Cost of Carbon", 
	                 "Unmet Demand", 
	                 "Value of a Statistical Life",
	                 "Covid Rate"), 
	       range = upper-lower,
	       case = "Allow Crowding",
	       base= data[1, 1],                   # midpoint estimate
	       param_new =paste(case, ":", param)) # add parameter names and the range

tornado_data_C2 <- C2_bounds %>% 
	mutate(param = c("AV cost/mile", 
	                 "Social Cost of Carbon", 
	                 "Unmet Demand", 
	                 "Value of a Statistical Life", 
	                 "Covid Rate"), 
	       range = upper-lower,
	       case = "Extra Buses",
	       base= data[1, 2],                   # midpoint estimate
	       param_new =paste(case, ":", param)) # add parameter names and the range

tornado_data_C3 <- C3_bounds %>% 
	mutate(param = c("AV cost/mile", 
	                 "Social Cost of Carbon", 
	                 "Unmet Demand", 
	                 "Value of a Statistical Life", 
	                 "Covid Rate"), 
	       range = upper-lower,
	       case = "AV Dispatch",
	       base= data[1, 3],                   # midpoint estimate
	       param_new =paste(case, ":", param)) # add parameter names and the range

tornado_data_C4 <- C4_bounds %>% 
	mutate(param = c("AV cost/mile", 
	                 "Social Cost of Carbon", 
	                 "Unmet Demand", 
	                 "Value of a Statistical Life", 
	                 "Covid Rate"), 
	       range = upper-lower,
	       case = "Longer Buses",
	       base= data[1, 4],                   #midpoint
	       param_new =paste(case, ":", param)) # add parameter names and the range

# Recombine into one dataset
tornado_data <- rbind(tornado_data_C4, 
                      tornado_data_C3, 
                      tornado_data_C2, 
                      tornado_data_C1)

```
 Build the plot
 Credit goes to Stack Overflow user `kikoralston'
 See: https://stackoverflow.com/questions/55751978/tornado-both-sided-horizontal-bar-plot-in-r-with-chart-axes-crosses-at-a-given/55755671
 and Nicholas Christain

With modifications

### Reorder parameters
 
```{R}
order.parameters <- tornado_data %>%
	mutate(param_new=factor(x=param_new, levels=param_new)) %>%
  	select(param_new) %>% unlist() %>% levels()
cases <- str_split_fixed(order.parameters, " : ", 2)[ , 1]
variables <- str_split_fixed(order.parameters, " : ", 2)[ , 2]
# width of columns in plot (value between 0 and 1)
width <- 0.8

# get data frame in shape for ggplot and geom_rect
tornado_plot <- tornado_data %>% 
  	# gather columns Lower_Bound and Upper_Bound into a single column using gather
	gather(key='type', value='output.value', lower:upper) %>%
  # just reordering columns
	select(param, param_new, type, output.value, range, base, case) %>%
  # create the columns for geom_rect
	mutate(param_new=factor(param_new, levels=order.parameters),
         ymin=pmin(output.value, base),
         ymax=pmax(output.value, base),
         xmin=as.numeric(param_new)-width/2,
         xmax=as.numeric(param_new)+width/2,
	       ylim_u = Inf,
	       ylim_l = -Inf,
	       xlim_u = as.numeric(param_new)+0.5,
	       xlim_l = as.numeric(param_new)-0.5)

tornado_plot <- tornado_plot %>% mutate(param_type = paste(param, ":", type))

```

### create Tornado plot

```{r}
ggplot() + 
   geom_rect(data = tornado_plot,  
            aes(ymax=ylim_u, ymin=ylim_l, xmax=xlim_u, xmin=xlim_l, fill = case), 
            show.legend = FALSE, alpha=0.4) + guides(colour = FALSE) + 
  geom_rect(data = tornado_plot,  
            aes(ymax=ymax, ymin=ymin, xmax=xmax, xmin=xmin, fill=param_type)) + 
  scale_fill_manual(name = "Sensitivity", 
                    values=c("Unmet Demand : lower" = "orange2", 
                             "Covid Rate : lower" = "sienna2", 
                             "Value of a Statistical Life : lower" = "skyblue3", 
                             "Social Cost of Carbon : lower" = "darkolivegreen3", 
                             "AV cost/mile : lower" = "tomato3",
                             "Unmet Demand : upper" = "orange3", 
                             "Covid Rate : upper" = "sienna3", 
                             "Value of a Statistical Life : upper" = "skyblue4", 
                             "Social Cost of Carbon : upper" = "darkolivegreen4", 
                             "AV cost/mile : upper" = "tomato4", 
                             "Longer Buses" = "gray90", 
                             "Extra Buses" = "gray90", 
                             "Allow Crowding" = "gray80", 
                             "AV Dispatch" = "gray80"),  
                    breaks = c("Covid Rate : upper", 
                               "Value of a Statistical Life : upper", 
                               "Unmet Demand : upper", 
                               "Social Cost of Carbon : upper", 
                               "AV cost/mile : upper"), 
                    labels= c("Covid Rate", 
                              "Value of a Statistical Life", 
                              "Unmet Demand", 
                              "Social Cost of Carbon", 
                              "AV cost/mile"))+
  geom_rect(data = tornado_plot,  
            aes(ymax=base, ymin=base, xmax=xmax, xmin=xmin), color='black') + 
  theme_minimal() +
  theme(axis.title.y=element_blank(), 
        legend.position = 'right', 
        plot.background = element_rect(color = 'white'), 
        panel.grid.minor = element_blank(), 
        panel.grid.major.y = element_blank()) + 
  scale_x_continuous(name = 'Sensitivity', 
                     breaks = c(3, 8, 13, 18),
                     labels = unique(cases)) +
  scale_y_continuous(name="Total Cost", 
                     labels = dollar_format(), 
                     limits = c(20000000, 170000000)) + 
  coord_flip() + theme(axis.text.y.left = element_text(size = 14))

```

# Baseline Cost Stacked Bar Chart

```{R}
# Grab CSV of base case costs
base_case_costs <- read.csv("base_case_total_costs.csv") 

# rename policy alternatives from 1-4 to their names
names(base_case_costs) <- c('Cost', 'Allow Crowding', 'Extra Buses', 'AV Dispatch', 'Longer Buses')

# Expand variable names
base_case_costs$Cost <- recode(base_case_costs$Cost, 
                       'O&M' = "Operations & Maintenance", 
                       'SCC' = 'Social Cost of Carbon',
                       'VSL' = 'Cost of Covid-related Deaths',
                       'externalities' = 'Congestion & pollutants')

# rearrange for stacked bar format
ra.melt <- melt(base_case_costs)
ra.melt$Cost <- factor(ra.melt$Cost, c('Social Cost of Carbon', 
                                       'Congestion & pollutants', 
                                       'Cost of Covid-related Deaths', 
                                       'Operations & Maintenance'))
names(ra.melt) <- c('Source', 'Alternative', 'Cost')

# plot
ggplot(ra.melt, aes(x=Alternative, y=Cost, fill=Source)) + theme_bw() +
  scale_y_continuous(labels=dollar_format()) + 
  geom_bar(stat='identity') +
 scale_fill_manual(values=c("Operations & Maintenance" = "skyblue3", 
                            'Cost of Covid-related Deaths' = "sienna3", 
                            'Congestion & pollutants' = "goldenrod2", 
                            "Social Cost of Carbon" = "darkolivegreen")) +
  theme(panel.grid.major.x = element_blank())

```

# Bus risk by Passenger Count
```{r}
# Grab data
bus_risk <- read.csv("bus_risk.csv")

# rename columns
names(bus_risk) <- c('X', 'bus.risk.40ft', 'bus.risk.60ft')

# Find the risk at the passenger capacity limits for each bus
limit60 <- bus_risk[bus_risk['X']==25, 'bus.risk.60ft'] # passenger limit of 25
limit40 <- bus_risk[bus_risk['X']==15, 'bus.risk.40ft'] # passenger limit of 15

# plot
ggplot(bus_risk, show.legend = FALSE) + 
  geom_line(aes(x=X, y=bus.risk.60ft/X, colour = '60ft bus risk'), 
            linetype = 'dashed', lwd=1.3, show.legend = FALSE) +
  scale_y_continuous(labels=percent_format())+
  geom_line(aes(x=X, y=bus.risk.40ft/X, colour='40ft bus risk'), 
            lwd=1.3, linetype = 'dotdash') + 
  geom_point(x=25, y=limit60/25, aes(colour='Covid passenger limit'), 
             pch = 23, lwd=3, fill='black') + 
  geom_point(x=15, y=limit40/15, aes(colour='Covid passenger limit'), 
             lwd=3,shape = 23, fill = 'black')+
  scale_color_manual('Infection Risk', 
                     values = c("Covid passenger limit"="black", 
                                "60ft bus risk"="sienna2", 
                                "40ft bus risk" =  "skyblue2")) + 
  ylab('Percent Infection Risk for Passengers') + 
  xlab('Passengers on the bus')+
  theme_bw()

```

# Histograms of infections by Policy Alternative in the Base Case
```{R}
# Grab data
base_mc <- read.csv("monte_carlo_infections.csv")

# 4 x 1 subplots (vertical)
par(mfcol=c(4,1), mar= c(5,4,1,1)) 

# breaks => Freedman-Diaconis bucketing
C1 <- base_mc$infections.C1
C2 <- base_mc$infections.C2
C3 <- base_mc$infections.C3
C4 <- base_mc$infections.C4
hist1 <- hist(C1, breaks = seq(min(C1), max(C1), length.out = 17), 
              ylab = 'Simulations', 
              xlab = 'Number of Infections while Allowing Crowding', 
              main = " ", col='darkolivegreen', xlim = c(0, 200))
hist2 <- hist(C2,  breaks = seq(min(C2), max(C2), length.out = 17), 
              ylab = 'Simulations', 
              xlab = 'Number of Infections with Extra Buses', 
              main = " ", col='skyblue4', xlim = c(0, 200))
hist3 <- hist(C3,  breaks = seq(min(C3), max(C3), length.out = 17), 
              ylab = 'Simulations', 
              xlab = 'Number of Infections while Dispatching AVs', 
              main = " ", col='orange3', xlim = c(0, 200))
hist4 <- hist(C4,  breaks = seq(min(C4), max(C4), length.out = 17), 
              ylab = 'Simulations', 
              xlab = 'Number of Infections with Longer Buses', 
              main = " ", col='sienna4', xlim = c(0, 200))

```
# Monte Carlo Convergence Justification
```{r}
# Cumulative Average after each run
p_hat_1 <- cumsum(base_mc$infections.C1)/seq_along(C1)
p_hat_2 <- cumsum(base_mc$infections.C2)/seq_along(C2)
p_hat_3 <- cumsum(base_mc$infections.C3)/seq_along(C3)
p_hat_4 <- cumsum(base_mc$infections.C4)/seq_along(C4)
p_hat = data.frame(C1 = p_hat_1, C2 = p_hat_2, C3 = p_hat_3, C4 = p_hat_4)

# Y axis bounds
y_gap <- max(p_hat_1)-min(p_hat_1)

# 2x2 subplots
op <- par(mfcol=c(2,2), mar = c(4, 5, 0.5, 1) + 0.1)

# Allow Crowding
plot(seq_along(C1), p_hat_1, type = "l", xlab = "N", 
     ylab =expression(hat(p)), ylim = c(min(p_hat_1), min(p_hat_1)+y_gap))
legend('topright', 'Allow Crowding', bty='n')

# Extra Buses
plot(seq_along(C2), p_hat_2, type = "l", xlab = "N", 
     ylab =expression(hat(p)), ylim = c(min(p_hat_2), min(p_hat_2)+y_gap))
legend('topright', 'Extra Buses', bty='n')

# AVs
plot(seq_along(C3), p_hat_3, type = "l", xlab = "N", ylab =expression(hat(p)), 
     ylim = c(min(p_hat_3), min(p_hat_3)+y_gap))
legend('topright', 'Autonomous Vehicles', bty='n')

# Longer buses
plot(seq_along(C1), p_hat_4, type = "l", xlab = "N", ylab =expression(hat(p)), 
     ylim = c(min(p_hat_4), min(p_hat_4)+y_gap))
legend('topright', 'Longer Buses', bty='n')

# Comparison of the mean value at n=500 and n=1000 with the difference between
convergence_table <- data.frame('X500' = t(p_hat[500, ]), 'X1000' = t(p_hat[1000, ]))
convergence_table$delta <- convergence_table$X1000 - convergence_table$X500
kable(convergence_table)
```


# Table of Mean Infections
## mean infections for the base case for the policy alternatives including first degree of community infection.
```{r}
# reproduction rate + original case count (1)
r_rate <- 1.97

# direct infections
infections_summary <- data.frame(
  allow.crowding = c(min(C1), mean(C1), max(C1), mean(C1 > 100)),
  extra.buses = c(min(C2), mean(C2), max(C2), mean(C2 > 100)),
  AVs = c(min(C3), mean(C3), max(C3), mean(C3 > 100)),
  longer.buses = c(min(C4), mean(C4), max(C4), mean(C4 > 100)),
  #including first degree of infections
  allow.crowding.d1 = c(min(C1)*r_rate, mean(C1)*r_rate, max(C1)*r_rate, mean(C1*r_rate > 100)),
  extra.buses.d1 = c(min(C3)*r_rate, mean(C2)*r_rate, max(C2)*r_rate, mean(C2*r_rate > 100)),
 AVs.d1 =c(min(C3)*r_rate, mean(C3)*r_rate, max(C3)*r_rate, mean(C3*r_rate > 100)),
 longer.buses.d1 = c(min(C4)*r_rate, mean(C4)*r_rate, max(C4)*r_rate, mean(C4*r_rate > 100)), 
 row.names = c('min', 'mean', 'max', 'p(>100)'))

kable(round(t(infections_summary), 2))
```


```{r}
# January 1st 2020 to Mid February 2020
Jan_Feb <- read.csv('C:/Users/lilyr/Documents/EPP/Research/AV/January_February_trip_file.csv')
# convert time from string to date time
Jan_Feb$start_time <- as.POSIXct(strptime(Jan_Feb$trip_start_time,"%Y-%m-%dT%H:%M:%SZ"))
# sum the peak loads to the trip start minute
Jan_Feb_agg <- aggregate(Jan_Feb$trip_max_load, list(Jan_Feb$start_hour), sum)
# mid Apr 2020 to May 31st 2020
Apr_May <- read.csv('C:/Users/lilyr/Documents/EPP/Research/AV/April_May_trip_file.csv')
# convert time from string to date time
Apr_May$start_time <- as.POSIXct(strptime(Apr_May$trip_start_time,"%Y-%m-%dT%H:%M:%SZ"))
# sum the peak loads to the trip start minute
Apr_May_agg <- aggregate(Apr_May$trip_max_load, list(Apr_May$start_hour), sum)
```
# Hourly Aggregated peak loads pre and during the pandemic
```{r}
# Plot on same plot
lims = c(5, 22)
plot(Jan_Feb_agg$Group.1, Jan_Feb_agg$x, xlab = 'Hour', ylab='Total passengers', 
     col= 'skyblue3',lwd=2, type = "l", xlim=lims)
points(Apr_May_agg$Group.1, Apr_May_agg$x, col = 'sienna3', lwd=2, type = "l", xlim = lims)
legend('topright',legend= c('January-February', 'April-May'), col=c('skyblue3', 'sienna3'), lty = 1, lwd = 2)

```

# Trade Off Plot between mitigation costs and infections

```{r}
full_costs <- read.csv('full_costs.csv')
# Comparing Covid rate assumptions
#trade_offs <- data.frame(t(full_costs[c(1,2,3,4,13,38, 39, 40, 41), c('Unnamed..0.1', 'Base.Scenario', 'Double.Covid.rate')]))
# comparing upper bound of demand
trade_offs <- data.frame(t(full_costs[c(1,2,3,4,38,39,40,41), c('Unnamed..0.1', 'Base.Scenario', 'halfway.to.Pre.pandemic.ridership')]))

# formatting
colnames(trade_offs) <- trade_offs[1, ]
trade_offs <- trade_offs[-1, ]
trade_offs[1:10] <- lapply(trade_offs, as.numeric)
trade_offs$case <- row.names(trade_offs)

# plot
ggplot()+
  geom_line(data=trade_offs, aes(x=`infections C1`, y=`total D1 C1`), 
            color = 'skyblue3', lwd=1.5, linetype = 'dotted')+
  geom_line(data = trade_offs, aes(x=`infections C2`, y=`total D1 C2`), 
            color = 'goldenrod3',  lwd=1.5, linetype = 'dotdash')+
  geom_line(data = trade_offs, aes(x=`infections C3`, y=`total D1 C3`), 
            color = 'skyblue4',  lwd=1.5)+
  geom_line(data = trade_offs, aes(x=`infections C4`, y=`total D1 C4`), 
            color = 'sienna3', lwd=1.5, linetype = 'dashed') + 
  # end points
  geom_point(data=trade_offs, aes(x=`infections C1`, y=`total D1 C1`), pch=19, 
             color = 'skyblue3', lwd=4) +
  geom_point(data = trade_offs, aes(x=`infections C2`, y=`total D1 C2`), pch=19, 
             color = 'goldenrod3',  lwd=4)+
  geom_point(data = trade_offs, aes(x=`infections C3`, y=`total D1 C3`), pch=19, 
             color = 'slategray4',  lwd=4)+
  geom_point(data = trade_offs, aes(x=`infections C4`, y=`total D1 C4`), pch=19, 
             color = 'sienna3', lwd=4) + 
  ylab('Cost (Millions)') +
  xlab('Infections')+ 
  scale_y_continuous(labels=scales::dollar_format(scale=1e-6))+theme_minimal() 

```